hbm_paper_results

### 2024-01-11 paper_main_results.Rmd

**qc - find ppts with centiles at 1 or 0**

@@

combined_preds %>% 
  filter_at(vars(pheno_list.z), any_vars(. %in% c(NA, Inf, -Inf))) %>%
  dplyr::select(participant) %>%
  unique()

# test.df <- combined_preds %>% 
#   dplyr::select(pheno_list)
# 
# test.df[test.df==1] <- 0
# 
# format(round(max(do.call(pmax, test.df)), 20), nsmall = 20)
```

@@

p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-biases in centile score errors across combat configurations", 
               color = "black", size = 14, hjust=.65))
p_sexannotate
ggsave("figs/sex_diff_win_feat.jpeg", p_sexannotate, width=7, height=8)
```

@@

wmv_centile_fan <- ggplot(wmv.df) +
  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_years, y=ukb_cn_df$WMV), alpha=0.4, width=.2, color="lightgray") +
  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
  geom_line(aes(x=age_years, y=centile_1, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_2, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_3, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_4, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
  geom_line(aes(x=age_years, y=centile_6, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_7, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_8, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_9, color=dataset, linetype=dataset)) +
  facet_wrap(~sex) + 
  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
  theme_minimal() +
  xlab("Age (yrs)") + ylab("Volume (mm3)") +
  theme_linedraw(base_size = 11) +
  theme(plot.title = element_text(size=12)) +
  labs(title="WMV brain charts")
wmv_centile_fan
```

@@

Combine plots
```{r}
fig_1 <- ggarrange(
  wmv_centile_fan, global_cent_errs_plot,
  ncol=2, nrow=1, 
  widths = c(1.3, 1),
  #labels=c("A", "B"),
  #hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "bottom",
  labels="auto"
  )

@@

>>>>

**QC: find ppts with centiles at 1 or 0**

@@

combined_preds %>% 
  filter_at(vars(pheno_list.z), any_vars(. %in% c(NA, Inf, -Inf))) %>%
  dplyr::select(participant) %>%
  unique()
```

@@

p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-biases in centile score errors across combat configurations", 
               color = "black", size = 14, hjust=.65))
p_sexannotate
ggsave("figs/sex_diff_win_feat.jpeg", p_sexannotate, width=7, height=8)
saveRDS(p_sexannotate, file="figs/sex_diff_win_feat.RDS")
```

@@

wmv_centile_fan <- ggplot(wmv.df) +
  geom_jitter(data = ukb_cn_df, aes(x=ukb_cn_df$age_years, y=ukb_cn_df$WMV), alpha=0.4, width=.2, color="lightgray") +
  #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
  geom_line(aes(x=age_years, y=centile_1, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_2, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_3, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_4, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_5, color=dataset, linetype=dataset), linewidth=1.5) +
  geom_line(aes(x=age_years, y=centile_6, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_7, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_8, color=dataset, linetype=dataset)) +
  geom_line(aes(x=age_years, y=centile_9, color=dataset, linetype=dataset)) +
  facet_wrap(~sex) + 
  scale_linetype_manual(values = c('ukb-CN-simsites'= "dotted", cf.gam = "solid", cf.gamlss = "solid", cf.lm = "solid"), name = "ComBat \nConfiguration", labels=dataset_names) +
  #scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#E84617", cf.gamlss = "#17E846", cf.lm = "purple")) +
  scale_color_manual(values = c('ukb-CN-simsites'= "black", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) + 
  xlab("Age (yrs)") + ylab("Volume (mm3)") +
  theme_linedraw(base_size = 11) +
  theme(plot.title = element_text(size=12)) +
  labs(title="WMV brain charts")
wmv_centile_fan
```

@@

not sure whether to include
```{r}
#look at sex distributions of ppts across sites
data.p <- subj_sum.df %>%
  count(sim.site, sex) %>% 
  group_by(sim.site) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = sim.site, y = prop, fill = sex)) +
  geom_col() +
  theme_linedraw(base_size = 11) +
  ggtitle("Distributions of male and female subjects across simulated sites")
```

Combine plots

```{r}
fig_1 <- ggarrange(
  wmv_centile_fan, global_cent_errs_plot,
  ncol=2, nrow=1, 
  widths = c(1.3, 1),
  #labels=c("A", "B"),
  #hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "bottom",
  labels="auto"
  )

@@

## FIGURE 2
need to run `paper_prop_results.Rmd` to get panel B

```{r}
prop_sex_mean_cent.plot.abs <- readRDS(file="figs/mean_subj_cent_err.RDS")
```

```{r}
fig_2 <- ggarrange(
  p_sex, prop_sex_mean_cent.plot.abs,
  ncol=2, nrow=1, 
  widths = c(1, 1.2),
  #labels=c("A", "B"),
  #hjust=0.01,
  # align = "hv",
  common.legend = FALSE,
  labels="auto"
  )
# fig_2

fig_2_annotate <- annotate_figure(fig_2, top = text_grob("Figure 2. Sex-biases in centile score errors across combat configurations", color = "black", hjust = 0, x = 0))
fig_2_annotate
ggsave("figs/figure_2.jpeg", fig_2_annotate, width=14, height=6)
```

### 2024-01-11 paper_prop_results.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, ggh4x)
#Sys.setenv("MC_CORES"=10L)
options("mc.cores")

source("./cubic_scripts/R_scripts/stats_tests.R")
aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
```

@@

```{r}
n_prop_list <- c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F")
```

@@

Read in predicted centiles & z scores
```{r}
pred.csvs <- list.files(path = paste0(data_path, "/ukb_ratio"), pattern = "_predictions.csv", full.names = TRUE)
combined_preds <- data.frame()

@@

  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))

    # Bind the data to the combined dataframe
    df_list <- c(df_list, list(data))
    assign("ratio_pred_list", df_list)
}
dim(ratio_pred_list)

#merge into single df
ratio_pred.df <- bind_rows(ratio_pred_list)
```

@@

```{r}
dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")

prop_sex_mean_cent.plot <- ratio_subj.df %>%
  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
  ggplot(aes(x=sim.site, y=mean_cent_diff*100, fill=sex)) +
  geom_violin() +
  #geom_smooth(method="lm") +
  theme_linedraw(base_size=12) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_grid(dataset ~ prop, labeller = labeller(dataset = as_labeller(dataset_names), prop=as_labeller(percent_f))) +
  xlab("Site") + ylab("Mean Centile Error")
prop_sex_mean_cent.plot
```

@@

```{r, warning=FALSE}
raw_centiles <- combined_preds %>%
  dplyr::filter(dataset == "ukb_CN")

merge.df <- full_join(raw_centiles, diffs.df)
unique(merge.df$dataset)
dim(raw_centiles)
dim(merge.df)

merge.df <- merge.df %>%
  dplyr::select(!File_ID)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, ggh4x, purrr)
#Sys.setenv("MC_CORES"=10L)
options("mc.cores")

source("./cubic_scripts/R_scripts/stats_tests.R")
#aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
```

@@

```{r}
dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
n_prop_list <- c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F")
```

@@

Read in predicted centiles & z scores FROM RAW DATA ONLY (to save memory)
```{r}
pred.csvs <- list.files(path = paste0(data_path, "/ukb_ratio"), pattern = "raw_predictions.csv", full.names = TRUE)

@@

  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = gsub("_data|_predictions.csv|prop-|[0-9]|-", "", Source_File),
           prop=gsub("-raw_predictions.csv", "", Source_File))

    # Bind the data to the combined dataframe
    df_list <- c(df_list, list(data))
    assign("ratio_raw_pred_list", df_list)
}
length(ratio_raw_pred_list)

#merge into single df
ratio_raw_pred.df <- bind_rows(ratio_raw_pred_list)
dim(ratio_raw_pred.df)
```

@@

```{r}
prop_sex_mean_cent.plot <- ratio_subj.df %>%
  dplyr::filter(dataset == "cf.gam" | dataset=="cf.gamlss") %>%
  ggplot(aes(x=sim.site, y=mean_cent_diff*100, fill=sex)) +
  geom_violin() +
  #geom_smooth(method="lm") +
  theme_linedraw(base_size=12) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_grid(dataset ~ prop, labeller = labeller(dataset = as_labeller(dataset_names), prop=as_labeller(percent_f))) +
  xlab("Site") + ylab("Mean Centile Error")
prop_sex_mean_cent.plot
```

@@

```{r, message=FALSE}
ratio.merge.df <- full_join(ratio_raw_pred.df, ratio.df)
unique(ratio.merge.df$dataset)
dim(ratio_raw_pred.df)
dim(ratio.merge.df)

ratio.merge.df <- ratio.merge.df %>%
  dplyr::select(!File_ID) %>%
  mutate(dataset=gsub("_predictions.csv","", dataset))
```

split back into separate dataframes for each proportion

```{r}
# df_list <- c()
# for (p in unique(ratio.merge.df$prop)) {
#   df <- ratio.merge.df %>%
#     dplyr::filter(prop == p)
#   df_list[p] <- list(df)
# }
# length(df_list) == length(unique(ratio.merge.df$prop))
```

define function to remove extremes
```{r}
rm_ext_cents <- function(og.data, feat_list){
  df_ext_list <- c()
  count <- 1
  for(pheno in feat_list) {
    list <- c(pheno, paste0("diff_", pheno), paste0("abs.diff_", pheno))

    df <- og.data %>%
      dplyr::select(participant, sex, age_days, sim.site, Source_File, dataset, prop, all_of(list)) %>% #get correct cols
      group_by(participant) %>% 
      dplyr::filter(!any((get(pheno) < 0.05 | get(pheno) > 0.95) & dataset == "raw")) %>% #drop ppts if raw centile is <0.05 or >0.95
      ungroup() %>%
      dplyr::filter(dataset != "raw") #now drop raw vals
    df_ext_list[count] <- list(df)
    count <- count + 1
  }
  stopifnot(length(df_ext_list) == length(pheno_list))
  df.no_ext <- Reduce(full_join, df_ext_list)
  return(df.no_ext)
}
```

apply fun

```{r}
ratio_no_ext_list <- mclapply(df_list, rm_ext_cents, feat_list=pheno_list, mc.preschedule = FALSE)
```
